Delayed effects of climate on vital rates lead to demographic divergence in Amazonian forest fragments

### 2021-08-18 review/Response to reviews.Rmd

> I suggest that a little more about the limitations of this modeling approach in the discussion would help set the context for interpreting the results and make the paper more robust.
> In some ways it seems like the analysis is adding more precision (by using monthly climate data) than the plants perceive and that would be measured in a yearly census.
> While the authors mention this, what about adding something short about when using this type of analysis is critical and when it might not be (should all demographers adopt this approach? Why or why not?)

##  {#R1-6}

> I was curious about why the paper did not include analyses of recruitment.
> Perhaps it is beyond the scope of this paper, but I wondered how new recruits responded to drought and wet also, both in number, as well as how big they were in the first year.

The reviewer is correct that recruitment is an important aspect of demography that could be effected by SPEI and fragmentation.
However, we believed the analyses in this paper were complicated enough to not include additional analyses.
We intend to follow up this paper with analyses of recruitment and integrating all vital rate models with an integral projection model.
We simply felt doing all of this (describing and interpreting DLNMs, modeling recruitment, and integrating all into an IPM) would be too much for a single paper.

@@

# Reviewer 3

**General comments**

> This is a really nice addition to the growing literature on lagged climate effects on plant demography, and also expands the tool kit for these studies.
> The combination of experimentally generated fragmentation with long term demographic data is also unique and very valuable.
> The paper is very clear and concise, with a few minor exceptions discussed in detail below.
> The one surprising omission was that the authors didn't use an Integral Projection Model to put the vital rates together into a metric of population growth potential.
> I enjoyed reading this.

We thought it would be too much to describe and interpret DLNMs for the vital rates, model recruitment, *and* integrate all those models into an integral projection model.
We though the description and interpretation of the DLNMs for survival, size, and flowering were complex and interesting enough to stand on their own.
We do intend to follow up with an IPM in a future publication.
See also our response to \@ref(R1-6).

### 2021-08-23 review/Response to reviews.Rmd

<!--# yes, but might change number of knots, so might require more work revising this -->

##  {#R3-5}

> Line 205-207 -- "number of knots not important ..." yes, but package mgcv includes tools for checking those assumptions, would be helpful to include residual analysis and the results of gam.check() in an appendix.

This comment was especially helpful and made us re-think limiting the number of "knots" to 3 in the SPEI dimension.
The term *k* is probably more accurately described as the "maximum basis complexity" or "maximum basis dimension".
We had thought that by limiting k to 3, we were simply limiting the complexity of the smooth in the SPEI dimension.
However, after taking the reviewer's suggestions we realized that there were other unexpected consequences.
We increased *k* for the SPEI dimension for all models and for other smooths for some models.
This resulted in some changes in the the shape of the crossbasis smooth and which areas of it were statistically significant.
We have updated the methods and results sections accordingly.
<!--# TODO: update results section! -->\
\
We have also now included a supplemental document with model validation, including checks that *k* is adequately large
.

### 2021-08-23 README.Rmd
closes #92 (but does not update results yet)

>>>>

**Using HPC**:
Also, you may need to install packages on the remote HPC if using `tar_make_clustermq()` with the SSH option.  Unfortunately, I'm not sure if there is an easy way to do this using `renv`. You should, however, make sure the same versions of the packages are installed locally and on the HPC, especially `gratia` which I'm using a development version of that has important bugs fixed, but not new bugs introduced.

### 2021-08-23 doc/paper.Rmd
closes #92 (but does not update results yet)

where the crossbasis function, $f \cdot w (d,l)$, is composed of two marginal basis functions: the standard predictor-response function $f(d)$, and the additional lag-response function $w(l)$.
These marginal functions are combined as a tensor product smooth such that the shape of one marginal function varies smoothly along the other dimension (see chapter 5 of @woodGeneralizedAdditiveModels2017 and @gasparriniPenalizedFrameworkDistributed2017 for more detail).
Penalized cubic regression splines were used for both marginal bases of the crossbasis function, with 35 knots for the lag dimension (i.e. number of lagged SPEI values for each observation with 36 months as a maximum lag) and 3 knots for the drought response dimension to restrict the shape of the fitted response to drought to bimodal when most complex.
Because of penalization, the number of knots is generally not important as long it is large enough to allow the smooth to represent the 'true' relationship [@woodGeneralizedAdditiveModels2017].
<!--# reference supplemental model diagnostics here --> Estimated degrees of freedom (edf) represent the 'true' complexity of the smooth after penalization with edf = 1 being equivalent to a straight line and larger numbers representing more complex curves.

>>>>

where the crossbasis function, $f \cdot w (d,l)$, is composed of two marginal basis functions: the standard predictor-response function $f(d)$, and the additional lag-response function $w(l)$.
These marginal functions are combined as a tensor product smooth such that the shape of one marginal function varies smoothly along the other dimension (see chapter 5 of @woodGeneralizedAdditiveModels2017 and @gasparriniPenalizedFrameworkDistributed2017 for more detail). Penalized cubic regression splines were used for both marginal bases of the crossbasis function. We increased *k*, the maximum basis dimension, as necessary to be sure it was adequate. Because of penalization, the maximum basis dimension is generally not important as long it is large enough to allow the smooth to represent the 'true' relationship [@woodGeneralizedAdditiveModels2017].
<!--# reference supplemental model diagnostics here --> Estimated degrees of freedom (edf) represent the 'true' complexity of the smooth after penalization with edf = 1 being equivalent to a straight line and larger numbers representing more complex curves.
